---
title: "Introduction to R- Coding Challenge 1"
Subtitle: Amanda Matioli
output:
  html_document:
    df_print: paged
---

This assignment is intended to introduce you to R, RStudio, and integration with GitHub, as well as help you practice some R coding. You may work in pairs. When submitting the assignment on Canvas, please put both names on the assignment. 

1.	5 pts. Explain the following things about R and Rstudio: 
o	Where do you type code? 

A- In the R source portion, i can do that by opening a new R script file. I can also type on the console portion, but is not recommended, once in the script portion i can do any modifications.

o	Where is the output of the code? 

A- In the console portion.

o	Where do plots show up? 

A- In the plot section in the right botton portion of the terminal

o	What is a global environment? 

A- Is the environment where all the objects and setted data are storage.

o	How do you get help for a function through R studio? 

A- You can go to the windows "help" (usually in the right connor of the R) or type help(name of the source)

o	What is an R package? 

A- They are a collection of functions and tools that helps extending R capabilities.

o	What is a function? 

A- Its like a mechanism to calculate any object that we may want to. Its a code that performance a specific task.

o	How do you find the installed and loaded packages? 

A- In the "Packages' portion of the R (The same place as the Plot and Help portion). Or i can just type installed.packages().

o	What is a working directory, and how do you find it? 

A- Its the place that you are extracting your input files and also the place that will be your outputs and save environment.
You can find it by using:
```{r}
getwd()
```

o	What is a relative file path and how is it different than an absolute file path? 

A- The relative file path would be the directory or subdirectory that you are at that moment, without the full description as your home directory. Its better for reproducibility.


2.	2 pts. Explain the steps to start a new R project connected to a GitHub repository.

A- You first open a new project and went to tools, in the global options you set the Git/SVN for the path in your Git executable portion and add my SSH key from my GitHub repository.

3.	3 pts. Please explain the different data classes and how they are similar and different from each other. 

o	Vector
o	Dataframe
o	Matrix

A- An vector is an collection of values of the same data type (numeric, character, etc)
B- A dataframe is a table that can have different data types in different columns.
C- A matrix is a structure where all values of the same data type are arranged in rows and columns.

Similarities: They both store data; matrices and dataframes are two-dimensional while vectors are one-dimensional; matrices need to have the same data type (numerical, character) while dataframes allows differents data types.

4.	10 pts. Complete all tasks in an R script and push it to GitHub.

o	Create a vector named 'z' with the values 1 to 200
```{r}
z <- c(1:200)
```

o	Print the mean and standard deviation of z on the console
```{r}
mean(z)
sd(z)
```
o	Create a logical vector named zlog that is 'TRUE' for z values greater than 30 and 'FALSE' otherwise.
```{r}
zlog <- z > 30

zlog
```

o	Make a dataframe with z and zlog as columns. Name the dataframe zdf
```{r}
zdf <- data.frame(z, zlog)

```

o	Change the column names in your new dataframe to equal “zvec” and “zlogic”
```{r}
colnames(zdf) <- c("zvec", "zlogic")
```

o	Make a new column in your dataframe equal to zvec squared (i.e., z2). Call the new column zsquared. 
```{r}
zdf$zsquared = z^2

zdf
```
o	Subset the dataframe with and without the subset() function to only include values of zsquared greater than 10 and less than 100 
```{r}
zdf2 <- subset(zdf, zsquared > 10 & zsquared < 100)
zdf.nosubset <- zdf[zdf$zsquared > 10 & zdf$zsquared < 100,]
```

o	Subset the zdf dataframe to only include the values on row 26
```{r}
zdf.26 <- subset(zdf, row(zdf) == 26)
```

o	Subset the zdf dataframe to only include the values in the column zsquared in the 180th row.
```{r}
zdf.180 <- subset(zdf, row(zdf) == 180, select = zsquared)
# zdf[180, "zsquared"] easy way to find
```
```{r}
getwd()
```
5.	3 pts. Download the Tips.csv file from Canvas. Use the read.csv() function to read the data into R using a relative file path  and so that the missing values are properly coded. **Note the missing values are reported in the data as a period (i.e., “.”). How do you know the data were read correctly? 

```{r}
Tips <- read.csv("TipsR.csv", na.strings = ".")
head(Tips)
```

A- To identify if the data were read correctly i use the command "head" to look into the file. I also look at the length of the file to check if it is the same as the original and i can also press to open the file in the environment pane and look trough it.

6.	2 pts. Annotate your code, commit the changes and push it to your GitHub repository you created last class
```{r}
# https://github.com/Matioli-amoc/PLPA6820-Amanda.git
```

