---
title: "Loops and functions annotation"
author: "Amanda Chaves"
date: "2026-03-17"
output: html_document
---

str() # Function that allow us to see the structure of the data

Function to convert Fahrenheit to Celsius:

```{r}
(5*(f_temp - 32)/9)
(5*(32 - 32)/9)

# If you want to do more convertions, you need to copy and paste the equation and modify the 
#degree_f portion into the number of the fahrenheit. We can skip this turning it into a function.

F_to_C <- function(f_temp){
  celsius <- (5*(f_temp - 32)/9)
  return(celsius)
}

F_to_C(32)
F_to_C(80)

```
We first start with naming the function something using the backwards arrow. Then we type “function()”.

After opening a new function we type the curly brackets. We will type the code we want to perform within the curly brackets.

Inside the parentheses after function is where we put our variables separated by commas.

sample.function <- function(... variable goes here ...){
  .... code goes here.... 
  return(... output ...)
}


# Now you try
Write a function to convert celsius to fahrenheiht using the conversion of F = C x (9/5)+32

```{r}
C_to_F <- function(c_temp){
  fahrenheit <- (c_temp * (9/5)+32)
  return(fahrenheit)
}

C_to_F(30)
C_to_F(-1)

```
# Iteration functions in base R

```{r}
 #repetition function
rep("A", 3)
rep(c("A", "B"), 10)
rep(c(1,2,3), 4, each = 5)

#sequence of numbers
#1:7
seq(from = 1, to = 7)
seq(from = 0, to = 10, by = 2)

#Allows you to generate a sequence of numbers based on a non-integer value (character vector).
# LETTERS
seq_along(LETTERS)

```

# For Loop

```{r}
for (i in 1:10){
  print(i*2)
}
```

We can combine functions and iterations

```{r}
for (i in -30:100){
  result <- F_to_C(i)
  print(result)
}
```

How we can save an object from a result of a loop?

```{r}

celsius.df <- NULL
for (i in -30:100){
  result <- data.frame(F_to_C(i), i)
  celsius.df <- rbind.data.frame(celsius.df, result)
}


```


A more practical example involving fungicide sensitivity data:

```{r}
library(tidyverse)
install.packages("drc")
library(drc)
EC50.data <- read.csv("EC50_all.csv")

isolate1 <- drm(100* EC50.data$relgrowth[EC50.data$is == "ILSO_5-41c"] ~
                  EC50.data$conc[EC50.data$is == "ILSO_5-41c"],
                fct = LL.4(fixed = c(NA, NA, NA, NA),
                           names = c("Slope", "Lower", "Upper", "EC50")),
                na.action = na.omit)

#output the summary of paramters including the estimate, standard error, t-value, and p-value outputs it into a data frame called summary.mef.fit for 'summary fit'

summary.fit <- data.frame(summary(isolate1)[[3]])

#outputs the summary of just the EC50 data including the estimate, standard error, upper and lower bounds of the 95% confidence intervals around th eEC50

EC50 <- ED(isolate1, respLev = c(50), type = "relative",
           interval = "delta")[[1]]


```
How i turn it into a situation that i loop the isolate code?

```{r}
nm <- unique(EC50.data$is)
nm

EC50.LL4 <- NULL

for (i in seq_along(nm)) {
  isolate1 <- drm(100* EC50.data$relgrowth[EC50.data$is == nm[[i]]] ~
                  EC50.data$conc[EC50.data$is == nm[[i]]],
                fct = LL.4(fixed = c(NA, NA, NA, NA),
                           names = c("Slope", "Lower", "Upper", "EC50")),
                na.action = na.omit)
  
  print(nm[[i]])
  summary.fit <- data.frame(summary(isolate1)[[3]])
  
  EC50 <- ED(isolate1, respLev = c(50), type = "relative",
           interval = "delta")[[1]]
  isolate.ec_1 <- data.frame(nm[[i]], EC50)
  EC50.LL4 <- rbind.data.frame(EC50.LL4, isolate.ec_1)
  EC50
  
}




```

Another way to do that is using tidyverse:

```{r}
EC50.data %>%
  group_by(is) %>%
  nest() %>%
  mutate(ll.4.mod = map(data, ~drm(.$relgrowth ~ .$conc,
                                   fct = LL.4(fixed = c(NA, NA, NA, NA),
                                              names = c("Slope", "Lower", "Upper", "EC50"))))) %>%
  mutate(ec50 = map(ll.4.mod, ~ED(.,
                                  respLev = c(50),
                                  type = "relative",
                                  interval = "delta")[[1]])) %>%
  unnest(ec50)




```




