---
title: "Code challenging VI"
author: "Amanda Chaves"
date: "2026-03-25"
output:
  md_document:
  html_document:
---

[GitHub link](https://github.com/Matioli-amoc/PLPA6820-Amanda.git)

1.	1 pts. Regarding reproducibility, what is the main point of writing your own functions and iterations? 

A- Using functions and iterations, regarding reproducibility, we can turn the code more efficient and scalable, allowing us to reuse the functions in different datasets and this can also reduce the human errors.

2.	2 pts. Conceptual: In your own words, describe: 1) how to write a function in R, and 2) how to write a for loop in R. In your answer, be specific about: basic syntax, where the code is written, and how results are returned or stored. (Answer in plain text; no code required for this question.)

A- To writa a function you use the "function" tool, first you set the keyword that are going to be your function name, after that you use the arrow to set (<-) and use the function tool followed by parenthesis where you list the arguments and variables for your function. The steps that the function are going to do are written inside curly brackets, where you write the code using the input to get the results. To get the results we use the command "return".
   To write a for loop in R you need to use the word "for" followed by one variable, inside parenthesis, that will be the values from the sequence/list (usually uses 'i'). The action that you want to repeat are going to be written inside curly brackets, and the code will run once for each value in sequence. The results for the loop usually goes to an variable setted inside the loop, and in the end we can print that variable to see the result.

This dataset contains the population and coordinates (latitude and longitude) of the 40 most populous cities in the US, along with Auburn, AL. Your task is to create a function that calculates the distance between Auburn and each other city using the Haversine formula. To do this, you'll write a for loop that goes through each city in the dataset and computes the distance from Auburn. Detailed steps are provided below.

3.	1 pts. Read in the Cities.csv file from Canvas using a relative file path.

```{r}
library(tidyverse)
cities <- read.csv("Cities.csv")

```



4.	6 pts. Function writing: Write a function that calculates the distance (in kilometers) between two pairs of geographic coordinates using the Haversine formula. 
a.	Function requirements:
i.	Inputs: lat1, lon1, lat2, lon2
ii.	The function must return a single numeric value called distance_km
iii.	All code shown below must be placed inside the function body

Use the following code inside your function:

convert to radians
rad.lat1 <- lat1 * pi/180
rad.lon1 <- lon1 * pi/180
rad.lat2 <- lat2 * pi/180
rad.lon2 <- lon2 * pi/180

Haversine formula
delta_lat <- rad.lat2 - rad.lat1
delta_lon <- rad.lon2 - rad.lon1
a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
c <- 2 * asin(sqrt(a)) 

Earth's radius in kilometers
earth_radius <- 6378137

Calculate the distance
distance_km <- (earth_radius * c)/1000

```{r}
h_distance <- function(lat1, lon1, lat2, lon2) {
  
  # convert to radians
rad.lat1 <- lat1 * pi/180
rad.lon1 <- lon1 * pi/180
rad.lat2 <- lat2 * pi/180
rad.lon2 <- lon2 * pi/180

# Haversine formula
delta_lat <- rad.lat2 - rad.lat1
delta_lon <- rad.lon2 - rad.lon1
a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
c <- 2 * asin(sqrt(a))

# Earth's radius in kilometers
earth_radius <- 6378137

# Calculate the distance
distance_km <- (earth_radius * c)/1000

return(distance_km)

}
```


5.	6 pts. Using your function from question 4, calculate the distance between Auburn, AL and New York City.
a.	Subset or extract only the latitude and longitude values needed from Cities.csv
b.	Pass those values into your function
c.	Store the result in an object

The expected output is 1367.854 km (small rounding differences are acceptable).

```{r}
auburn <- cities[cities$city == "Auburn",]
ny <- cities[cities$city == "New York",]

dist_result <- h_distance(
  lat1 = auburn$lat,
  lon1 = auburn$long,
  lat2 = ny$lat,
  lon2 = ny$long
)

dist_result

```


6.	4 pts. Use your distance function inside a for loop to calculate the distance between Auburn, AL and every other city in Cities.csv. Each iteration should calculate the distance for one city

The first 9 distances printed should match the values shown below
## [1] 1367.854
## [1] 3051.838
## [1] 1045.521
## [1] 916.4138
## [1] 993.0298
## [1] 1056.022
## [1] 1239.973
## [1] 162.5121
## [1] 1036.99

```{r}
distances <- NULL

for (i in seq_along(cities$city)) {
  dist <- h_distance(
  lat1 = auburn$lat,
  lon1 = auburn$long,
  lat2 = cities$lat[[i]],
  lon2 = cities$long[[i]]
  )
  
print(dist)
distances <- c(distances, dist)
  
}



```


7.	4 pts. Modify your loop so that each iteration appends one new row to a dataframe with the following columns:

•	City1 (the other city)
•	City2 (Auburn)
•	Distance_km

The dataframe should grow inside the loop, and the first six rows should look like this:
##         City1  City2 Distance_km
## 1    New York Auburn   1367.8540
## 2 Los Angeles Auburn   3051.8382
## 3     Chicago Auburn   1045.5213
## 4       Miami Auburn    916.4138
## 5     Houston Auburn    993.0298
## 6      Dallas Auburn   1056.0217

```{r}

df_dist <- NULL

for (i in seq_along(cities$city)) {
  dist <- h_distance(
  lat1 = auburn$lat,
  lon1 = auburn$long,
  lat2 = cities$lat[[i]],
  lon2 = cities$long[[i]]
  )
  
  new_row <- data.frame(
    City1 = cities$city[[i]],
    City2 = "Auburn",
    Distance_Km = dist
  )
  
df_dist <- rbind.data.frame(df_dist, new_row)
  
}

head(df_dist)


```


8.	1 pts. Commit and push a gfm .md file to GitHub inside a directory called Coding Challenge 6. Provide me a link to your github written as a clickable link in your rendered .pdf, .docx, or .html
