---
title: "Code challenging VII"
author: "Amanda Chaves"
date: "2026-03-26"
output:
  md_document:
  html_document:
---

[GitHub link!!!!](https://github.com/Matioli-amoc/PLPA6820-Amanda.git)

1.	4 pts. Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor

STAND$Treatment <- as.factor(STAND$Treatment) # example shown here. 

```{r}
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)
PlantE <- read.csv("PlantEmergence.csv")

PlantE$Treatment <- as.factor(PlantE$Treatment)
PlantE$DaysAfterPlanting <- as.factor(PlantE$DaysAfterPlanting)
PlantE$Rep <- as.factor(PlantE$Rep)

```


2.	5 pts. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. 

```{r}
lmp <- lm(Emergence ~ Treatment*DaysAfterPlanting, data = PlantE) 
summary(lmp)
```

```{r}
anova(lmp)

```


3.	5 pts. Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.


A- No, we dont need the interaction term, because p value is > than 0.05.

```{r}
lm_model <- lm(Emergence ~ Treatment + DaysAfterPlanting, data = PlantE)
summary(lm_model)

```

```{r}
anova(lm_model)
```

A- The intercept represents the mean for Treatment 1 at the first day. When we look at the results for Treatment 2, we can see that, on average, the treatment 2 results in 134.53 fewer comparing with treatment 1. That shows us that the difference between treatments is significant (p<0.001)

4.	5 pts. Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.

```{r}
lmp_lsmeans <- emmeans(lm_model, ~ Treatment)
result_lmpmeans <- cld(lmp_lsmeans, alpha = 0.05, details = TRUE)
result_lmpmeans
```


5.	4 pts. The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 

```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}
 
plot_cldbars_onefactor(lm_model, "Treatment")

```

A- The bars are representing the mean of the merged plans for each treatment, and the error bars shows the standard error. The letters shows the significant differences, for the Tukey test at alpha = 0.05. The treatments that share the same letter are not significantly different, while the treatments that shows different letters are significantly different. For example: The treatment 2 has significantly fewer emerged plants than most of the other treatments, while the treatments 1,3,4,5,6,8 and 9 do not show any significant difference among each other. The treatment 7, the letter "b" shows that this treatment is significantly lower than the treatments labeled "a" (3, 5 and 6) but does not differ form treatments labeled "ab"

6.	2 pts. Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the .md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to your github.
